% #############################################################################
% Agradecimentos / Acknowledgments
% !TEX root = ../main.tex
% #############################################################################

\noindent I would like to thank Professor Rodrigo
Rodrigues, not only for his supervision of this thesis but for
over four years of mentoring and countless doors he has opened for me.
Without a doubt, my academic, professional and
personal development would not have been the same without his
intervention. I would also like to thank Professor Peter
Druschel, who welcomed me multiple times at the Max Planck
Institute for Software Systems in Saarbrücken, an institution
where I learned a lot of lessons which I will carry with me and
for his astute comments and scientific rigour, from which I have
learned a lot.

I would like to acknowledge several other colleagues, mentors and
contributors, who have been with me through my
academic journey: Afonso Tinoco, Roberta De Viti, Professor
Deepak Garg, Professor Vijay Chidambaram, Professor Bobby
Bhattacharjee and Professor Pedro Adão.

I would like to thank my parents, for the housing and food they
have provided me through the years, an indispensible prerequisite for good
research. Also, for raising me.

Finally, I would like to thank my friends. Although words are not
nearly enough to describe the shared moments (both of joy and
hardship), the companionship, the comraderie and the bond that
connects us, I am forever grateful: Afonso, Andreia, Chico,
Cintrão, Donga, Filipa, Gaspar, Homem, João, Lamego, Lé, Maria, Mariana, Marta Ambrósio,
Marta Sousa, Pacheco, Pedro, Pina, Regouga, Reynolds, Rocha,
Sanches, Vasco, Zé e Leonor.
